Nested sampling with two objective functions

### 2015-10-26 Paper/ms.tex
Random work

The prior $\pi(\xx)$ implies a certain prior for $L_1$ and $L_2$, which we
denote $\pi(L_1, L_2)$. An example of a prior is shown as the
probability density in Figure~\ref{fig:joint}. The partition function
$Z(\beta_1, \beta_2)$ is a set of expected values with respect to this density.
Simple Monte Carlo sampling from $\pi$ will not work except for values of
$(\beta_1, \beta_2)$ where the canonical distribution remains similar to $\pi$.
We need a sampler that explores regions where $L_1$ and $L_2$ are high in order
to accurately estimate $Z(\beta_1, \beta_2)$.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/joint.pdf}
\caption{\it The particle at the corner of the grey rectangle
has a corner count of 1, since 1 other particle is within the
grey rectangle.\label{fig:joint}}
\end{figure}

@@

\section{Proposed algorithm}

@@

The corner count of particle $i$ is defined as the number of particles with
a lower value of {\em both} objective functions compared to particle $i$.
Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
particles in the rectangle whose upper right corner is at
$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint} for an
example). The equation for the corner count is:
\begin{eqnarray}
C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
L_2(\xx_j) < L_2(\xx_i)
\right).\label{eqn:corner_count}
\end{eqnarray}

\section{What kind of accuracy can we expect?}
Imagine computing $Z(\beta_1, \beta_2)$ by looping over a grid of $\beta_2$
values and doing an NS run with respect to $L_1$ for each. For each
value of $\beta_1$ and $\beta_2$ that we're interested in, we can expect
a (marginal) ``one-sigma'' accuracy in
$\ln\left[Z(\beta_1, \beta_2)\right]$ of $\pm \sqrt{H(\beta_1, \beta_2)/N}$,
for a computational cost proportional to...

>>>>

The prior $\pi(\xx)$ implies a certain prior for $L_1$ and $L_2$, which we
denote $\pi(L_1, L_2)$. An example of a prior is shown as the
probability density in Figure~\ref{fig:joint1}. The partition function
$Z(\beta_1, \beta_2)$ is a set of expected values with respect to this density.
Simple Monte Carlo sampling from $\pi$ will not work except for values of
$(\beta_1, \beta_2)$ where the canonical distribution remains similar to $\pi$.
We need a sampler that explores regions where $L_1$ and $L_2$ are high in order
to accurately estimate $Z(\beta_1, \beta_2)$.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/joint1.pdf}
\caption{\it An example of a prior for $L_1$ and $L_2$, which is implied by
the prior $\pi(\xx)$. We have overplotted some points drawn from this prior.
\label{fig:joint1}}
\end{figure}

Our proposed algorithm (introduced in Section~\ref{sec:algorithm}) makes use
of a quantity we call the corner count. For any values $(\ell_1, \ell_2)$,
the prior mass for which both objective functions are lower is
\begin{eqnarray}
X(\ell_1, \ell_2) &=& \int \pi(L_1, L_2) \mathds{1}\left(L_1 < \ell_1, L_2 < \ell_2 \right)
 \, dL_1 \, dL_2
\end{eqnarray}
This is the two dimensional cumulative distribution function (CDF) of
$L_1$ and $L_2$.

The corner count of particle $i$ is defined as the number of particles with
a lower value of {\em both} objective functions compared to particle $i$.
Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
particles in the rectangle whose upper right corner is at
$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
example). The equation for the corner count is:
\begin{eqnarray}
C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
L_2(\xx_j) < L_2(\xx_i)
\right).\label{eqn:corner_count}
\end{eqnarray}

@@

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/joint2.pdf}
\caption{\it The same as Figure~\ref{fig:joint1}.
\label{fig:joint2}}
\end{figure}

\section{Proposed algorithm}\label{sec:algorithm}

@@

\section{What kind of accuracy can we expect?}

Imagine computing $Z(\beta_1, \beta_2)$ by looping over a grid of $\beta_2$
values and doing an NS run with respect to $L_1$ for each. For each
value of $\beta_1$ and $\beta_2$ that we're interested in, we can expect
a (marginal) ``one-sigma'' accuracy in
$\ln\left[Z(\beta_1, \beta_2)\right]$ of $\pm \sqrt{H(\beta_1, \beta_2)/N}$,
for a computational cost proportional to...

\section{What kind of accuracy do we need?}

Use Skilling's argument about the prior distribution of a log Bayes factor.
Even though the performance of this algorithm won't be great for high $H$, it
will be good enough. I think the probability it won't give enough accuracy will
be ``exponentially small''.
